---
title: "BasicTutorial"
output: html_document
date: "2025-04-03"
---


## Installing clootl

clootl is currently in review at CRAN. For the time being you can install it from Github using devtools

```{r}
install.packages("devtools")
library(devtools)  
install_github("eliotmiller/clootl")
library(clootl)
```

## Getting a tree

The most recent version of the tree is available with the package.
You can easily get a subtree using "extractTree" and a list of scientific names.
This defaults tree version 1.4 and to the names in the 2023 taxonomy.

If you don't give it any arguments, it will return the whole tree (~11,000 tips)

```{r}

ex1 <- extractTree(species=c("Turdus migratorius",
                             "Setophaga dominica",
                             "Setophaga ruticilla",
                             "Sitta canadensis"))

plot(ex1)
```

You should cite the author's whose work went into generating these relationships whenever possible. You can get the citations to the input phylogenies using 'getCitations'.

```{r}

cites <- getCitations(ex1)
cites

```

To get the tree using a different taxonomy year, use the argument "taxonomy_year".
Current options are 2021, 2022. And 2024 is coming soon.

```{r}

ex2 <- extractTree(species=c("Turdus migratorius",
                             "Setophaga dominica",
                             "Setophaga ruticilla",
                             "Sitta canadensis"),
                             taxonomy_year = 2021)

```

You can also use eBird codes as tip labels,.
```{r}
ex3 <- extractTree(species=c("amerob",
                             "canwar",
                             "reevir1",
                             "yerwar",
                             "gockin"), 
                              label_type="code")
plot(ex3)
```
