---
title: "AvonetTutorial"
output: html_document
date: "2025-04-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing clootl

clootl is currently in review at CRAN. For the time being you can install it from Github using devtools. We also need a few other packages for this demo

```{r install,  include=FALSE}
library(devtools)
install_github("eliotmiller/clootl")
install.packages("phytools")
install.packages("readxl")
library("phytools")
```

## Getting avonet data

An important feature of our new phylogeny is its interoperability and connection to existing databases and resources. As an example of this feature, we illustrate how the phylogeny can be connected to AVONET [@tobias_avonet_2022], a database of morphological measurements for the world's birds.

To access AVONET, use the `readxl` [@wickham2019package] package to access the database directly from Figshare.

```{r}
library(readxl)

# Define the file URL and destination
file_url <- "https://figshare.com/ndownloader/files/34480856"
destfile <- tempfile(fileext = ".xlsx")

# Download the file
download.file(file_url, destfile, mode = "wb")

# Read the sheet that corresponds to eBird taxonomy
dat <- as.data.frame(read_excel(destfile, sheet = "AVONET2_eBird"))

# Create a column with underscores for simplicity later
dat$underscores <- sub(" ", "_", dat$Species2)
```


Importantly, AVONET was published in the 2021 version of the eBird/Clements taxonomy [@clements_ebird/clements_2019]. While AVONET does contain Avibase taxon IDs [@lepage_avibase_2014] that can be used to fast-forward the database to newer taxonomies–and we strongly recommend doing this for real research use–the simplest approach, for example purposes, is to extract a phylogeny in 2021 taxonomy.

```{r}
# Take a random sample of 50 species
spp <- sample(dat$Species2, 50)

# Extract a tree for these species
prunedTree <- extractTree(species=spp, label_type="scientific", taxonomy_year=2021, version="1.4")

```

This leaves us with a paired dataset and phylogeny. Align these further, and use phytools [@revell2024phytools] to visualize the distribution of body mass across the phylogeny of these 50 random taxa.

``{r}
library(phytools)
prunedDat <- dat[dat$Species2 %in% spp,]

# Pull a vector of traits out, log transform for normality
x <- log(prunedDat$Mass)
names(x) <- prunedDat$underscores

# Plot log body mass across the phylogeny
contMap(prunedTree, x, fsize=0.5)
```