---
title: "BasicTutorial"
output: html_document
date: "2025-04-03"
---


## Install clootl


```{r}
library(devtools)  
install_github("eliotmiller/clootl")
library(clootl)
```

## Getting the data repo

The most recent version of the tree (currently 1.4) is available with the package.
If you want other versions, you need to use the GitHub repo https://github.com/McTavishLab/AvesData
If you have already downloaded the Aves Data Repo to your computer,
you can pass in the path to that directory to get other tree versions
The data repo also contains the 100 tree dated sample sets. 

```{r, eval=FALSE}
ex_1_2 <- extractTree(species=c("Turdus migratorius","Setophaga dominica", "Setophaga ruticilla", "Sitta canadensis"),
                               version=1.2,
                               data_path="~/projects/otapi/AvesData")
```


Alternately, you can automatically download all the data from the AvesData github repo.
It will be unpacked as folder in whatever directory you pass in. 
This example creates a folder in your current working directory with the data named "AvesData-main".
It will also save the location to your R environment file, so that clootl automatically knows where to look for the data.
```{r}
get_avesdata_repo(path=".") 
```
Then you can use extractTree to get otehr versions of the tree.

```{r}

ex2_1_3 <- extractTree(species=c("Turdus migratorius",
                             "Setophaga dominica",
                             "Setophaga ruticilla",
                             "Sitta canadensis"),
                             version=1.3)

plot(ex2_1_3)
```

